---
title: "Titanic Survival Analysis"
output: html_notebook
---
***
**Load the titanic clean data created as part of data wrangling excercise into titanic dataframe.**

```{r, echo=TRUE}
# 0: Load the data in RStudio
titanicOrginal <- read.csv("titanic_clean.csv")
titanic <- select (titanicOrginal, pclass, survived, sex, age, has_cabin_number)
titanic <- filter(titanic, !is.na(pclass) )
titanic <- mutate(titanic, survived = ifelse(survived == 1, "Survived", "Not Survived"), pclass = factor(pclass), has_cabin_number = factor(ifelse(has_cabin_number == 1, "Has Cabin Number", "No Cabin Number")))
```
*NA from pclass is filtered out and meaningful names are given to axis labels*

***

**Plot the total counts of male and female across class**

```{r, echo=TRUE}
ggplot(titanic, aes(fill = sex, x = pclass)) +
  geom_bar(position = "dodge")
```

***

**Plot the survival numbers across sex and class**

```{r, echo=TRUE}
ggplot(titanic, aes(fill = sex, x = pclass)) +
  geom_bar(position = "dodge") + 
  facet_grid(". ~ survived")
```

***

**Plot the survival number as scatter plot using Jitter with survival status in seperate frames**

```{r, echo=TRUE}
# Position jitter (use below)
posn.j <- position_jitter(0.5, 0)

# Use ggplot() for the last instruction
ggplot(titanic, aes(col = sex, x = pclass, y = age)) +
  geom_jitter(size = 3, alpha = .5, position = posn.j) + 
  facet_grid(". ~ survived")
```
*The above plot clearly shows survival rate is more in females*

***

**Plot the survival numbers with gender in seperate frames**

```{r, echo=TRUE}
ggplot(titanic, aes(col = survived, x = pclass, y = age)) +
  geom_jitter(size = 3, alpha = .5, position = posn.j) + 
  facet_grid(". ~ sex")
```
*This clearly shows passengers in 3rd class irrespective of gender didnt get much chance at survival compared to first class.*

***

**Does cabin number got any significance on survival? data wrangling question**

```{r, echo=TRUE}
ggplot(titanic, aes(col = survived, x = pclass, y = age)) +
  geom_jitter(size = 3, alpha = .5, position = posn.j) + 
  facet_grid(". ~ has_cabin_number")
```
*Cabin number is mostly allocated to first class passengers and hence it does appear to have a significance in survival rate*

***

